ScPy: A SuperCollider Extension for Performing Numerical Computation via Embedded Python

### 2016-08-28 doc/paper.tex
extras

ScPy is developed and maintained as an open-source project under the GPLv3 license. The latest code
and documentation is available on GitHub\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}.

@@

@@

@@

def fn(x):
    x = pv.forward(x)
    x = pv.to_bin_offset(x)
    x.imag = -x.imag
    x = pv.from_bin_offset(x)
    x = pv.backward(x)
    return x
    \end{minted}
    \caption{An effect which inverts the frequency of each analysis bin across the center of
    the bin, creating an unusual detuning effect.}
    \label{lst:binflipper}
\end{listing}

@@

One issue involves the process of passing data between SuperCollider and C++. Since UGens are
designed to work primarily with data streams, adding support for passing in many types of data came
with a number of challenges. Essentially, all data must be serialized into an array of floats.
Strings must be converted to an array of ASCII character codes, and prepended with their length.
Arguments passed through to the Python code can have one of many types, and therefore type
information must be encoded as well. However, just passing the type of a variable is not enough to
know what we can do with it, since we also want to know if it inherits from a class we can use.

@@

>>>>

ScPy is developed and maintained as an open-source project under the GPLv3 license. The latest
code, documentation, and issues are available on
GitHub\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}. Contributions to the project are
welcome.

@@

The choice of SuperCollider was largely because it is one of the most widely used audio programming
environments. Using SuperCollider also enabled the reuse of many parts of its codebase. Since it
already has STFT based operations, it was a good target for extension purposes in that area.

@@

Complex numbers are used extensively, even to represent non-Cartesian data such as polar
coordinates or phase vocoder bins in ScPy. In polar form, the real and imaginary parts represent
magnitude and phase, respectively. Similarly for the phase vocoder, real and imaginary parts
represent magnitude and frequency. This is done to take advantage of the limited data types
available in NumPy and relative convenience of working with complex numbers as opposed to tuples or
classes. Future work could extend NumPy to give more descriptive names to these data fields.

@@

def fn(x):
    x = pv.forward(x)
    x = pv.to_bin_offset(x)
    x.imag = -x.imag # imaginary part holds frequency
    x = pv.from_bin_offset(x)
    x = pv.backward(x)
    return x
    \end{minted}
    \caption{An effect which inverts the frequency of each analysis bin across the center of
    the bin, creating an unusual detuning effect.}
    \label{lst:binflipper}
\end{listing}

@@

One issue involves the process of passing data between SuperCollider and C++. Since UGens are
designed to work primarily with data streams, adding support for passing in many types of data came
with a number of challenges. Essentially, all data must be serialized into an array of floats.
Strings must be converted to an array of ASCII character codes, and prepended with their length.
Arguments passed through to the Python code can have one of many types, and therefore type
information must be encoded as well. However, just passing the type of a variable is not enough to
know what we can do with it, since we also want to know if it inherits from a class we can use. The
encoding and decoding implementation is unfortunately fairly complex, but creating a UGen with
inputs this complicated was likely not considered during the design of SuperCollider, so some
friction was to be expected in this area.

@@

Throughout~\autoref{sec:lit}, articles are discussed which contain processing techniques that could
be used in ScPy. Although implementations of many of these techniques are likely to become a part
of ScPy in the future, there is no obstacle to the reader implementing them as Python libraries
usable by ScPy or even as open-source contributions to the project.
